---
title: "STA303-Project-Part-2"
author: "Amrinder Sehmbi"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Imported Libraries
```{r, eval=TRUE, echo = T}
#oo <- options(repos = "https://cran.r-project.org/")
#install.packages("Matrix")
#install.packages("lme4")
#options(oo)


library(ggplot2)
library(knitr)
library(vcd)
library(lme4) 
library(influence.ME)
library(pROC)
library(lattice)
library(mlmhelpr)
```


# Load the data set in R
```{r, eval=TRUE, echo = T}
# Open CSV file with data
FullData <- read.csv("~/STA303/shotdetail_2022.csv")

# Remove ID columns
RemovedColumns <- FullData[, -c(which(names(FullData) %in% 
                                        c('GRID_TYPE', 'GAME_ID', 
                                          'GAME_EVENT_ID', 'PLAYER_ID',
                                          'TEAM_ID')))]
# Filter for Denver Nuggets Data
DenverNuggetsData <-subset(RemovedColumns, TEAM_NAME == "Denver Nuggets")

# Add column indicating if shot taken at home or away
DenverNuggetsData$HOME_AWAY <- ifelse(DenverNuggetsData$HTM == "DEN", "Home", 
                                      "Away")

```



# Data Summary
```{r, eval=TRUE, echo = T}
# Summary of Data
summary(DenverNuggetsData)

# Visual 1: Distribution of EVENT_TYPE
barplot(table(DenverNuggetsData$EVENT_TYPE), 
        main = "Shots Made vs Missed",
        xlab = "Shot Result",
        ylab = "Frequency",
        col = c("coral2", "cadetblue2"))

# Visual 2: Shot Zone vs Shot Outcome
barplot(table(DenverNuggetsData$EVENT_TYPE, DenverNuggetsData$SHOT_ZONE_BASIC), 
        beside = TRUE, 
        legend.text = c("Made", "Missed"),
        main = "Shot Outcome vs Shot Zone",
        xlab = "Shot Zone",
        ylab = "Count",
        col = c("coral2", "cadetblue2"),
        cex.axis = 1.5)

# Visual 3: Shots Made by Shot Type and Player Shooting
ShotsMadeData = subset(DenverNuggetsData, EVENT_TYPE == "Made Shot")
desired_categories <- c("Aaron Gordon", "Jamal Murray", "Nikola Jokic")
ShotsMadeData = subset(ShotsMadeData, PLAYER_NAME %in% desired_categories)
desired_categories <- c("Dunk Shot", "Jump Shot", "Layup Shot")
ShotsMadeData = subset(ShotsMadeData, ACTION_TYPE %in% desired_categories)

mosaicplot(table(ShotsMadeData$PLAYER_NAME, ShotsMadeData$ACTION_TYPE),
           main = "Shot Made by Player and Action Type",
           xlab = "Player",
           ylab = "Action Type",
           color = TRUE,
           las = 1,
           cex.axis = 1)  

```


# Model Fitting

## Model 1: Initial Model with all Variables
```{r, eval=TRUE, echo = T}
# Model 1: Initial Model with all Variables
model1 <- glmer(SHOT_MADE_FLAG ~ SHOT_ZONE_BASIC + SHOT_DISTANCE 
                + MINUTES_REMAINING + factor(HOME_AWAY) + factor(PERIOD) 
                + (1|ACTION_TYPE)+(1|ACTION_TYPE:PLAYER_NAME), 
                family = "binomial", data = DenverNuggetsData,
                control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
summary(model1)
mlmhelpr::icc(model1)
auc(roc(DenverNuggetsData$SHOT_MADE_FLAG, predict(model1, type = "response")))


```


## Model 2: Excluding MINUTES_REMAINING
```{r, eval=TRUE, echo = T}

# Model 2: Excluding MINUTES_REMAINING
model2 <- glmer(SHOT_MADE_FLAG ~ SHOT_ZONE_BASIC + SHOT_DISTANCE 
                + factor(HOME_AWAY) + factor(PERIOD) 
                + (1|ACTION_TYPE)+(1|ACTION_TYPE:PLAYER_NAME), 
                family = "binomial", data = DenverNuggetsData,
                control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))



summary(model2)
mlmhelpr::icc(model2)
auc(roc(DenverNuggetsData$SHOT_MADE_FLAG, predict(model2, type = "response")))
```

## Model 3: Excluding MINUTES_REMAINING, HOME_AWAY
```{r, eval=TRUE, echo = T}
# Model 3: Excluding MINUTES_REMAINING, HOME_AWAY
model3 <- glmer(SHOT_MADE_FLAG ~ SHOT_ZONE_BASIC + SHOT_DISTANCE + factor(PERIOD) 
                + (1|ACTION_TYPE)+(1|ACTION_TYPE:PLAYER_NAME), 
                family = "binomial", data = DenverNuggetsData,
                control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))



summary(model3)
mlmhelpr::icc(model3)
auc(roc(DenverNuggetsData$SHOT_MADE_FLAG, predict(model3, type = "response")))
```


## Model 4: Excluding MINUTES_REMAINING, HOME_AWAY, SHOT_DISTANCE
```{r, eval=TRUE, echo = T}
# Model 4: Excluding MINUTES_REMAINING, HOME_AWAY, SHOT_DISTANCE
model4 <- glmer(SHOT_MADE_FLAG ~ SHOT_ZONE_BASIC  + factor(PERIOD) 
                + (1|ACTION_TYPE)+(1|ACTION_TYPE:PLAYER_NAME), 
                family = "binomial", data = DenverNuggetsData,
                control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))



summary(model4)
mlmhelpr::icc(model4)
auc(roc(DenverNuggetsData$SHOT_MADE_FLAG, predict(model4, type = "response")))
```


## Model 5: Final Model excluding MINUTES_REMAINING, HOME_AWAY, SHOT_DISTANCE, PERIOD
```{r, eval=TRUE, echo = T}
# Model 5: Final Model
final.model <- glmer(SHOT_MADE_FLAG ~ SHOT_ZONE_BASIC 
                + (1|ACTION_TYPE)+(1|ACTION_TYPE:PLAYER_NAME), 
                family = "binomial", data = DenverNuggetsData,
                control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))



summary(final.model)
mlmhelpr::icc(final.model)
auc(roc(DenverNuggetsData$SHOT_MADE_FLAG, predict(final.model, type = "response")))

```


# Model Diagnostics

## Dfbetas
```{r, eval=TRUE, echo = T,fig.height=10}

infl <- influence(final.model,"ACTION_TYPE")

plot(infl,which="dfbetas",parameters=c(1),xlab="DFbetaS",ylab="Action Type")
plot(infl,which="dfbetas",parameters=c(2),xlab="DFbetaS",ylab="Action Type")
plot(infl,which="dfbetas",parameters=c(3),xlab="DFbetaS",ylab="Action Type")
plot(infl,which="dfbetas",parameters=c(4),xlab="DFbetaS",ylab="Action Type")
plot(infl,which="dfbetas",parameters=c(5),xlab="DFbetaS",ylab="Action Type")
plot(infl,which="dfbetas",parameters=c(6),xlab="DFbetaS",ylab="Action Type")
plot(infl,which="dfbetas",parameters=c(7),xlab="DFbetaS",ylab="Action Type")

```


## Cooks Distance
```{r, eval=TRUE, echo = T,fig.height=8}

plot(infl, which="cook", cutoff=4/nrow(ranef(final.model)$ACTION_TYPE), sort=TRUE, xlab="Cook´s Distance",ylab="Action type")
```

## Removing Outliers for ACTION_TYPE
```{r, eval=TRUE, echo = T}
# removing influential groups from data
remove_categories <- c("Jump Shot", "Running Jump Shot")
sub = subset(DenverNuggetsData, !(ACTION_TYPE %in% remove_categories))

# fit final model without influential observations
final.model.removed.outliers <- glmer(SHOT_MADE_FLAG ~ SHOT_ZONE_BASIC 
                + (1|ACTION_TYPE)+(1|ACTION_TYPE:PLAYER_NAME), 
                family = "binomial", data = sub,
                control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

summary(final.model.removed.outliers)
mlmhelpr::icc(final.model.removed.outliers)
auc(roc(sub$SHOT_MADE_FLAG,  predict(final.model.removed.outliers, type = "response")))
```


















